A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-08-09 text/manuscript.Rmd
Move figures to end of paper. Move dicussion in front of method. Order is now: intro, result, discussion, methods, main figs, supplemental figs

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi=600)
suppressPackageStartupMessages(c(
  library(rstan),
  library(foreach),
  library(sqldf),
  library(tidyverse),
  library(ggthemes),
  library(data.table),
  library(biomaRt),
  library(dplyr),
  library(GenomicRanges),
  library(SummarizedExperiment),
  library(reactome.db),
  library(plyr),
  library(cluster),
  library(pheatmap),
  library(clusterProfiler),
  library(org.Hs.eg.db),
  library(dbscan),
  library(class),
  library(TCGAbiolinks),
  library(patchwork),
  library(ggplotify),
  library(ComplexHeatmap),
  library(knitr),
  library(kableExtra),
  library(UpSetR),
  library(cowplot),
  library(doRNG),
  library(uwot),
  library(janitor),
  library(FactoMineR),
  library(seriation),
  library(rtracklayer),
  library(poibin),
  library(flextable)
))
```

@@

```{r child='../code/figure1.Rmd'}
```
In order to classify tumors using this dataset, we used multiple correspondence
analysis (MCA) [@JSSv025i01]. First, we determined the number of dimensions
containing useful information. (**figures 1B, C**). Then we performed a UMAP
analysis, both in order to summarize the MCA graphically, and as a preprocessing
step to boost the performance of density based clustering. The resulting map was
notable for its lobed structure, with several large, robust projections
reproducible regardless of random seed setting. A representative version of this
3D UMAP is shown in **figure 1D** as a 2D projection, rotated to enhance the
visibility of the major features. Following this spatial mapping we attempted to
define groupings of similar tumors within the spatial map using HDBSCAN, which
performs hierarchical clustering and provides metrics of cluster stability and
probabilities of cluster membership for each node. However, HDBSCAN is sensitive
to several parameters; key for our analysis are the minimum number of tumor
samples in a cluster that capture the maximum number of tumor samples, measured
by probability of membership of $\geq 5\%$ in at least one class. Thus, we
created a score metric as the fraction of classified tumors with max probability
$< 5\%$ in one class and chose a cluster size of 92 to minimize the score
function (**figure 1E**). HDBSCAN with these settings resulted in ten distinct
high-density clusters which we then projected onto the UMAP (**figure 1F**).
This classified 6,038 out of 7,607 tumors but still resulted in a
significant fraction of unclassified tumor samples. Since we ultimately wish to
be able to classify any tumor using this scheme, we performed k Nearest
Neighbors (kNN) analysis, which computes a similarity metric to every tumor in
the set and then lets the $k$ most similar tumors "vote" as to the identity of
the query tumor sample based on their classification labels. We set $k$ to be
the square root of the number of tumor samples (87). Using this method, we
assigned cluster membership to the remaining tumors (**figure 1G**).

@@

```{r child='../code/figure2.Rmd'}

```

@@

```{r child = '../code/figure4.Rmd'}

```

@@

```{r child='../code/table2.Rmd'}
```
<mark>Statement does not apply anymore. The highest # of enriched path is 3. Not many are in 
metastasis. Only 2. “Fc epsilon receptor (FCERI) signaling” and "Signaling by SCF-KIT" </mark>
<strike>There were also a number of pathways that were enriched with lower specificity
(among 4 or more clusters) which were found to be enriched in metastases. We
identified 18 such pathways of which 12 had significantly higher mutation rates
in cluster-specific metastases, including “G1/S DNA Damage Checkpoints”,
“Deubiquitination”, “Sumoylation”, “M Phase”, “RHO GTPase Effectors”, “TCF
dependent signaling in response to WNT”, “Class I MHC mediated antigen
processing & presentation”, “Chaperonin-mediated protein folding”, “Intrinsic
Pathway for Apoptosis”, “Cellular Senescence”, “DNA Double-Strand Break
Response”, and “Pre-NOTCH Expression and Processing” (**Table 1**). </strike>

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi=600, warning = FALSE)
suppressPackageStartupMessages(c(
  library(rstan),
  library(foreach),
  library(sqldf),
  library(tidyverse),
  library(ggthemes),
  library(data.table),
  library(biomaRt),
  library(dplyr),
  library(GenomicRanges),
  library(SummarizedExperiment),
  library(reactome.db),
  library(plyr),
  library(cluster),
  library(pheatmap),
  library(clusterProfiler),
  library(org.Hs.eg.db),
  library(dbscan),
  library(class),
  library(TCGAbiolinks),
  library(patchwork),
  library(ggplotify),
  library(ComplexHeatmap),
  library(knitr),
  library(kableExtra),
  library(UpSetR),
  library(cowplot),
  library(doRNG),
  library(uwot),
  library(janitor),
  library(FactoMineR),
  library(seriation),
  library(rtracklayer),
  library(poibin),
  library(flextable)
))
```

@@

In order to classify tumors using this dataset, we used multiple correspondence
analysis (MCA) [@JSSv025i01]. First, we determined the number of dimensions
containing useful information. (**figures 1B, C**). Then we performed a UMAP
analysis, both in order to summarize the MCA graphically, and as a preprocessing
step to boost the performance of density based clustering. The resulting map was
notable for its lobed structure, with several large, robust projections
reproducible regardless of random seed setting. A representative version of this
3D UMAP is shown in **figure 1D** as a 2D projection, rotated to enhance the
visibility of the major features. Following this spatial mapping we attempted to
define groupings of similar tumors within the spatial map using HDBSCAN, which
performs hierarchical clustering and provides metrics of cluster stability and
probabilities of cluster membership for each node. However, HDBSCAN is sensitive
to several parameters; key for our analysis are the minimum number of tumor
samples in a cluster that capture the maximum number of tumor samples, measured
by probability of membership of $\geq 5\%$ in at least one class. Thus, we
created a score metric as the fraction of classified tumors with max probability
$< 5\%$ in one class and chose a cluster size of 92 to minimize the score
function (**figure 1E**). HDBSCAN with these settings resulted in ten distinct
high-density clusters which we then projected onto the UMAP (**figure 1F**).
This classified 6,038 out of 7,607 tumors but still resulted in a
significant fraction of unclassified tumor samples. Since we ultimately wish to
be able to classify any tumor using this scheme, we performed k Nearest
Neighbors (kNN) analysis, which computes a similarity metric to every tumor in
the set and then lets the $k$ most similar tumors "vote" as to the identity of
the query tumor sample based on their classification labels. We set $k$ to be
the square root of the number of tumor samples (87). Using this method, we
assigned cluster membership to the remaining tumors (**figure 1G**).

@@

@@

@@

<mark>Statement does not apply anymore. The highest # of enriched path is 3. Not many are in 
metastasis. Only 2. “Fc epsilon receptor (FCERI) signaling” and "Signaling by SCF-KIT" </mark>
<strike>There were also a number of pathways that were enriched with lower specificity
(among 4 or more clusters) which were found to be enriched in metastases. We
identified 18 such pathways of which 12 had significantly higher mutation rates
in cluster-specific metastases, including “G1/S DNA Damage Checkpoints”,
“Deubiquitination”, “Sumoylation”, “M Phase”, “RHO GTPase Effectors”, “TCF
dependent signaling in response to WNT”, “Class I MHC mediated antigen
processing & presentation”, “Chaperonin-mediated protein folding”, “Intrinsic
Pathway for Apoptosis”, “Cellular Senescence”, “DNA Double-Strand Break
Response”, and “Pre-NOTCH Expression and Processing” (**Table 1**). </strike>

@@

```{r child='../code/figure1.Rmd'}
```

```{r child='../code/figure2.Rmd'}
```

```{r child = '../code/figure4.Rmd'}
```

```{r child='../code/table2.Rmd'}
```

![](../figures/fig6B.png)
![](../figures/fig6C.png)
